Variational Monte Carlo on bosonic systems

### 2018-03-25 project_1/article/article.tex
insignificant changes
Started on interacting particles result text

        Looking at the one-body part of the Hamiltonian we get
        \begin{align}
            H
            &=
            \sum_{k = 1}^N
            \brac{
                -\frac{\hbar^2}{2m}\nabla_k^2
                + \half m\brak{
                    \omega^2(x_k^2 + y_k^2)
                    + \omega_z^2 z_k^2
                }
            }
            \\
            &=
            \frac{\hbar\omega}{2}
            \sum_{k = 1}^N
            \brac{
                -{\nabla_k'}^2
                +
                \brak{
                    {x_k'}^2 + {y_k'}^2 + \lambda^2{z_k'}^2
                }
            },
        \end{align}
        where we have introduced the dimensionless frequency $\lambda =
        \omega_z/\omega$. The single particle functions also gets a scaling
        factor
        \begin{align}
            \phi_k
            &=
            \exp\brak{
                -\frac{\alpha\hbar}{m\omega}
                ({x_k'}^2 + {y_k'}^2 + \beta^2{z_k'}^2)
            }
            \equiv
            \phi(\vf{r}_k').
        \end{align}
        The correlation wavefunction and the interaction potential remain the
        same, but using the scaled hard core sphere radius $a/a_{\text{ho}} =
        0.0043$. For brevity we will remove the ticks on the variables.

@@

		For our special case : INSERT DERIVATION OF LOCAL ENERGY derivative wrt alpha??

@@

\subsection{Interacting elliptical harmonic oscillator}
         \begin{table}
             \caption{The expected energy, the standard
             deviation, the acceptance ratio and CPU time for each $alpha$
             for the elliptical potential with the interacting wavefunction. We
             have used importance sampling with $\delta t = 0.1$ and $2^{21}$
             Monte Carlo cycles and an additional $10\%$ of that as
             thermalization steps. The hard sphere radius is set to $0.0043$ and
             a $\beta = \lambda = \sqrt{8}$.}
             \centering
             \begin{ruledtabular}
                 \begin{tabular}{ccccc}
                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
                     &$A$&$t_C$ \\
                     \hline
                     0.2&35.17548&0.05399&0.98995&24.48868\\
                     0.3&27.62004&0.02311&0.98187&13.65132\\
                     0.4&24.97850&0.00827&0.97248&13.54666\\
                     0.5&24.39877&0.00030&0.96140&13.38388\\
                     0.6&24.83863&0.00604&0.94998&13.31570\\
                     0.7&25.82855&0.01059&0.93766&13.20579\\
                     0.8&27.22895&0.01436&0.92375&13.11933\\
                 \end{tabular}
             \end{ruledtabular}
             \label{tab:10_interacting}
         \end{table}

@@

>>>>

        Looking at the one-body part of the Hamiltonian we get
        \begin{align}
            H
            &=
            \sum_{k = 1}^N
            \brac{
                -\frac{\hbar^2}{2m}\nabla_k^2
                + \half m\brak{
                    \omega^2(x_k^2 + y_k^2)
                    + \omega_z^2 z_k^2
                }
            }
            \\
            &=
            \frac{\hbar\omega}{2}
            \sum_{k = 1}^N
            \brac{
                -{\nabla_k'}^2
                +
                \brak{
                    {x_k'}^2 + {y_k'}^2 + \lambda^2{z_k'}^2
                }
            },
            \label{eq:one_body_hamiltonian}
        \end{align}
        where we have introduced the dimensionless frequency $\lambda =
        \omega_z/\omega$. The single particle functions also gets a scaling
        factor
        \begin{align}
            \phi_k
            &=
            \exp\brak{
                -\frac{\alpha\hbar}{m\omega}
                ({x_k'}^2 + {y_k'}^2 + \beta^2{z_k'}^2)
            }
            \equiv
            \phi(\vf{r}_k').
        \end{align}
        The correlation wavefunction and the interaction potential remain the
        same, but using the scaled hard core sphere radius $a/a_{\text{ho}} =
        0.0043$. For brevity we will remove the ticks on the variables.

@@

		For our special case we only have one variational parameter, $\alpha$.  The expression for the derivative is, 
		\begin{equation}
			\frac{d E_L}{d\alpha} = 
				2\left( \expval{\frac{1}{\psi_T}\frac{d\psi_T}{d\alpha} E_L} - \expval{\frac{1}{\psi_T} \frac{d\psi_T}{d\alpha}}\expval{E_L} \right).
		\end{equation}

		INSERT DERIVATION OF LOCAL ENERGY??

@@

	\subsubsection{Test of Gradient descent}

	Before moving on to the interacting, more interesting systems, it would be beneficial to test the gradient descent implementation
	now that we are certain that it moves towards the optimum variational parameter $\alpha = \frac{1}{2}$.

\subsection{Interacting elliptical harmonic oscillator}

		We now make two changes to the simple systems in the previous section in order to make everything
		a bit more interesting. First, the potential in the Hamiltonian is perturbed in the $z$-direction according to
		\autoref{eq:one_body_hamiltonian} by setting $\lambda = \sqrt{8} \approx 2.82843$. This corresponds to enclosing particles
		of the simulation in a spherical trap instead of a circular trap. Second, a particle-to-particle interaction is 
		introduces in form of a hard shell with radius $\frac{a}{a_\text{ho}} = 0.0043$. Similarly to the non-interacting
		case, the simulation results for different values of $\alpha$ are displayed in \autoref{tab:10_interacting}.

		In \autoref{tab:10_interacting} the system is set to have $10$ interacting particles. Notice first that the energy is
		higher than for the interacting system than for a comparable non-interacting system (\autoref{tab:initial_optimum_comparison}).
		Moreover, the standard deviation gets lover as the system is close to an optimal variational parameter.  The importnance 
		sampling method is used in this simulations, with a relatively low time step $\delta t = 0.1$, leading to a correspondingly
		high acceptance ratio.

         \begin{table}
             \caption{The expected energy, the standard
             deviation, the acceptance ratio and CPU time for each $\alpha$
             for the elliptical potential with the interacting wavefunction. We
             have used importance sampling with $\delta t = 0.1$ and $2^{21}$
             Monte Carlo cycles and an additional $10\%$ of that as
             thermalization steps. The hard sphere radius is set to $0.0043$ and
             a $\beta = \lambda = \sqrt{8}$.}
             \centering
             \begin{ruledtabular}
                 \begin{tabular}{ccccc}
                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
                     &$A$&$t_C$ \\
                     \hline
                     0.2&35.17548&0.05399&0.98995&24.48868\\
                     0.3&27.62004&0.02311&0.98187&13.65132\\
                     0.4&24.97850&0.00827&0.97248&13.54666\\
                     0.5&24.39877&0.00030&0.96140&13.38388\\
                     0.6&24.83863&0.00604&0.94998&13.31570\\
                     0.7&25.82855&0.01059&0.93766&13.20579\\
                     0.8&27.22895&0.01436&0.92375&13.11933\\
                 \end{tabular}
             \end{ruledtabular}
             \label{tab:10_interacting}
         \end{table}

		The result of simulations with a higher number of particles, $n=50$ and $n=100$, are  shown in \autoref{tab:50_interacting}
		and \autoref{tab:100_interacting} respectively. We observe that the energy for systems of interacting particles increases
		rapidly as the size of the system increases.

@@

\section{Derivation of Local Energy alpha derivative}

TO BE INSERTED HERE??? Or whaaaat boiii...
